---
title: Traffic

---

```{r, echo=FALSE, include=FALSE}
# All defaults
## libraries
library(fs)
library(tibble)
library(dplyr)
library(stringr)
library(googlesheets4)
library(lubridate)
library(plotly)
```



<style type="text/css">
  .img-responsive {
      margin: 0 auto;
      max-width: 75%
  }
  
  .resp-container {
      position: relative;
      overflow: hidden;
      padding-top: 60.0%;
  }
  
  .resp-iframe {
      position: absolute;
      top: 0;
      left: 0;
      width: 100%;
      height: 150%;
      border: 0;
  }
      
  .Scorecard_subt {
    margin: 0px; 
    padding-bottom: 5px; 
    font-size: 16px; 
    text-align: left; 
    color:#7f7f7f
  }

@media screen and (min-width:1000px) {
  .flex-container {
    display: flex;
  }
  
  .Mycolumn {
    width: 50%;
    text-align: justify;
<!-- margin-right: 5px -->
  <!-- margin-left: 0px; -->
  <!-- background-color: #7f7f7f; -->
  }

  .Mycolumn2 {
    width: 50%;
    text-align: justify;
    <!-- margin-left: 0px; -->
    <!-- margin-right: 5px -->
  }

  .Scorecard {
    margin: 0px;
    padding: 0px;
    font-size: 32px;
    font-weight: bold;
    color: #161b21;
    text-align: left;
  }

  .Scorecard2 {
    margin: 0px;
    padding: 0px;
    padding-top: 7px;
    font-size: 26px;
    font-weight: bold;
    color: #161b21;
    text-align: left;
  }
}

@media screen and (max-width:1000px) {
  .Mycolumn {
    width: 100%;
    margin-right: 0px;
    text-align: justify;
  }
  
  .Mycolumn2 {
    width: 100%;
    margin-left: 0px;
    margin-right: 0px
    text-align: justify;
  }
  
  .Scorecard {
    margin: 0px;
    padding: 0px;
    font-size: 24px;
    font-weight: bold;
    color: #161b21;
    text-align: left;
  }

  .Scorecard2 {
    margin: 0px;
    padding: 0px;
    padding-top: 7px;
    font-size: 20px;
    font-weight: bold;
    color: #161b21;
    text-align: left;
  }
}

  </style>

The dashboard below provides an overview of the evolution of flights in the [EUROCONTROL Member State Area](/definition/eurocontrol-area/) (referred to as 'Network' below). A more detailed view of the traffic situation at network and stakeholder level is shown in our <a href="https://www.eurocontrol.int/Economics/DailyTrafficVariation-States.html " target="_blank">Daily traffic variation dashboard</a>. The results are shown in the dashboard below and also available for download in our [data repository](/data/).

For a view of the en-route and terminal [service units](/acronym/su/) used by the <a href="https://www.eurocontrol.int/crco" target="_blank">Central Route Charges Office of EUROCONTROL (CRCO)</a> to collect charges visit our <a href="https://www.eurocontrol.int/ServiceUnits/Dashboard/EnRouteMainDashboard.html" target="_blank">Service Units dashboard</a>.

And if you want to know more about forecasting products and services visit our [STATFOR Statistics and Forecasts page](statfor/).  


<!-- Network Traffic section -->
<div class="container" style="margin: 0px; padding: 0px">
#### Network View
This section shows traffic situation in terms of flights for the latest available day, rolling week and year to date with a comparison with relevant previous periods.
The chart presents the traffic evolution (7-day average number of flights) since 2019.

<!-- Daily traffic graphs and scorecards -->
<div class="flex-container" style=" margin-bottom: 5px;">
<div class="Mycolumn" >
<div class="row" style="margin:0px; margin-bottom: 5px;background-color: #f2f2f2; border-radius: 5px; width: 101%;">
<div class="col-12" ; style="margin: 0px;  ">
  <p id="last_day" style="font-weight: bold; margin-bottom: 0px;">Day</p>
</div>
<div class="col-3" ; style="margin: 0px; ">
  <p id="day_traffic" class="Scorecard"></p>
  <p class="Scorecard_subt">flights</p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
  <p id="dif_day_prev_week" class="Scorecard2" ></p>
  <p class="Scorecard_subt">vs prev. week</p>
</div>
<div class="col-3" ; style="margin: 0px; ">
  <p id="dif_day_2019" class="Scorecard2"></p>
  <p class="Scorecard_subt">vs 2019</p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
  <p id="dif_day_prev_year" class="Scorecard2"></p>
  <p class="Scorecard_subt">vs 2021</p>
</div>
</div>
<div class="row" style="margin:0px; margin-bottom: 5px;background-color: #f2f2f2; border-radius: 5px; width: 101%;">
<div class="col-12" ; style="margin: 0px;  ">
  <p id="last_week" style="font-weight: bold; margin-bottom: 0px;">Week</p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
  <p id="week_traffic" class="Scorecard"></p>
  <p id="avg_week_traffic" class="Scorecard_subt"></p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
  <p id="dif_week_prev_week" class="Scorecard2" ></p>
  <p class="Scorecard_subt">vs prev. week</p>
</div>
<div class="col-3" ; style="margin: 0px; ">
  <p id="dif_week_2019" class="Scorecard2"></p>
  <p class="Scorecard_subt">vs 2019</p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
  <p id="dif_week_prev_year" class="Scorecard2"></p>
  <p class="Scorecard_subt">vs 2021</p>
</div>
</div>
<div class="row" style="margin:0px; margin-bottom: 5px;background-color: #f2f2f2; border-radius: 5px; width: 101%;">
<div class="col-12" ; style="margin: 0px;  ">
  <p id="y2d_day_text" style="font-weight: bold; margin-bottom: 0px;">Year to date</p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
  <p id="y2d_traffic" class="Scorecard"></p>
  <p id="avg_y2d_traffic" class="Scorecard_subt"></p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
</div>
<div class="col-3" ; style="margin: 0px; ">
  <p id="dif_y2d_2019" class="Scorecard2"></p>
  <p class="Scorecard_subt">vs 2019</p>
</div>
<div class="col-3" ; style="margin: 0px;  ">
  <p id="dif_y2d_prev_year" class="Scorecard2"></p>
  <p class="Scorecard_subt">vs 2021</p>
</div>
</div>
</div>
<div class="Mycolumn2">
<div class="col-12" style=" margin: 0px; " >
<!-- <div class="resp-container"> -->
<!-- <iframe class="resp-iframe" src="https://datastudio.google.com/embed/reporting/4efb3da2-6cd8-4778-ab2b-0fd0cbe71db2/page/tvomC" frameborder="0" style="border:0" allowfullscreen></iframe> -->
<!-- </div> -->

<!-- script for traffic figures -->
<script type="text/javascript" src="https://www.gstatic.com/charts/loader.js"></script>

<script type="text/javascript">
  var Data
  var my_view_traffic
  var last_day
  var last_day_text
  var last_week_text

  var day_traffic
  var dif_day_prev_week
  var dif_day_2019
  var dif_day_prev_year

  var week_traffic
  var avg_week_traffic
  var dif_week_prev_week
  var dif_week_2019
  var dif_week_prev_year
 
  var y2d_traffic
  var y2d_prev_year_traffic
  var y2d_2019_traffic 
  var avg_y2d_traffic
  var dif_y2d_2019
  var dif_y2d_prev_year

  google.load("visualization", "1.1", {
    packages: ["corechart", "table", "controls", "bar"]
  });
  google.setOnLoadCallback(Initialisation);

  function Initialisation() {
    var query = new google.visualization.Query(
      "https://docs.google.com/spreadsheets/d/14Lt3C3PCxxYCLXXYxRrFXd1s0tVZ3EQ8q9YfuYvx3Ww/edit?usp=sharing&sheet=NW_Flights_All"
    );

  query.send(drawDashboard);
  }

  function drawDashboard(response) {
    Data = response.getDataTable();
    UpdateTraffic();
   }
//---------------------------Update traffic---------------------------------------------
  function UpdateTraffic() {
  my_view_traffic = new google.visualization.DataView(Data);

  //       for (var y = 1; y < my_view_traffic.getNumberOfColumns(); y++) {
  //        // my_view_traffic.getValue(x, y) 
  //        last_day = my_view_traffic.getValue(0, y)
  //      }
  last_day = my_view_traffic.getValue(0, 38); 
  my_view_traffic.setRows(
  Data.getFilteredRows([{ column: 7, value: last_day }])
  );

// Calculate datestrings
last_day_text = last_day.toDateString();
    last_day_text = last_day_text.substring(8, 11) + last_day_text.substring(4, 8) + last_day_text.substring(11)

  //calculate 1st day of rolling week and string
  var first_week_day = new Date();
    first_week_day.setDate(last_day.getDate()-6);
    first_week_day = first_week_day.toDateString();
  last_week_text = first_week_day.substring(8, 11) + (last_day_text.substring(3, 7) == first_week_day.substring(4, 8) ? "" : first_week_day.substring(4, 8)) + "- " + last_day_text;
  y2d_day_text = last_day_text.substring(3, 7) == "Jan " ? "01 - " + last_day_text : "01 Jan - " + last_day_text

// Calculate traffic scorecard values
   day_traffic = my_view_traffic.getValue(0, 11);
   dif_day_prev_week = Math.round(my_view_traffic.getValue(0, 18)*100);
     dif_day_prev_week = (dif_day_prev_week > 0) ? "+" + dif_day_prev_week : "" + dif_day_prev_week
   dif_day_2019 = Math.round(my_view_traffic.getValue(0, 20) * 100) ;
     dif_day_2019 = (dif_day_2019 > 0) ? "+" + dif_day_2019 : "" + dif_day_2019
   dif_day_prev_year = Math.round(my_view_traffic.getValue(0, 19) * 100);
    dif_day_prev_year = (dif_day_prev_year > 0) ? "+" + dif_day_prev_year : "" + dif_day_prev_year;


  week_traffic = my_view_traffic.getValue(0, 21)
  avg_week_traffic = Math.round(week_traffic/7)
  dif_week_prev_week = Math.round(my_view_traffic.getValue(0, 27)*100);
     dif_week_prev_week = (dif_week_prev_week > 0) ? "+" + dif_week_prev_week : "" + dif_week_prev_week
  dif_week_2019 = Math.round(my_view_traffic.getValue(0, 29)*100);
     dif_week_2019 = (dif_week_2019 > 0) ? "+" + dif_week_2019 : "" + dif_week_2019
  dif_week_prev_year = Math.round(my_view_traffic.getValue(0, 28)*100);
     dif_week_prev_year = (dif_week_prev_year > 0) ? "+" + dif_week_prev_year : "" + dif_week_prev_year

  y2d_traffic = my_view_traffic.getValue(0, 30);
  y2d_prev_year_traffic = my_view_traffic.getValue(0, 31);
  y2d_2019_traffic = my_view_traffic.getValue(0, 32);
  
  avg_y2d_traffic = Math.round(my_view_traffic.getValue(0, 33));
  dif_y2d_2019 = Math.round(my_view_traffic.getValue(0, 37)*100);
     dif_y2d_2019 = (dif_y2d_2019 > 0) ? "+" + dif_y2d_2019 : "" + dif_y2d_2019
  dif_y2d_prev_year = Math.round(my_view_traffic.getValue(0, 36)*100);
     dif_y2d_prev_year = (dif_y2d_prev_year > 0) ? "+" + dif_y2d_prev_year : "" + dif_y2d_prev_year

// Update scorecard values
  document.getElementById("last_day").innerHTML = "Day (" + last_day_text + ")";
  document.getElementById("last_week").innerHTML = "Rolling week (" + last_week_text + ")";
  document.getElementById("y2d_day_text").innerHTML = "Year to date (" + y2d_day_text + ")";

  <!-- document.getElementById("last_day_dly").innerHTML = "Day (" + last_day_text + ")"; -->
  <!-- document.getElementById("last_week_dly").innerHTML = "Rolling week (" + last_week_text + ")"; -->
  <!-- document.getElementById("last_day_dly_flt").innerHTML = "Day (" + last_day_text + ")"; -->
  <!-- document.getElementById("last_week_dly_flt").innerHTML = "Rolling week (" + last_week_text + ")"; -->
  
  document.getElementById("day_traffic").innerHTML = String(day_traffic).replace(/(.)(?=(\d{3})+$)/g, '$1,');
  document.getElementById("dif_day_prev_week").innerHTML = String(dif_day_prev_week) + '%';
  document.getElementById("dif_day_2019").innerHTML = String(dif_day_2019) + '%';
  document.getElementById("dif_day_prev_year").innerHTML = String(dif_day_prev_year) + '%';

  document.getElementById("week_traffic").innerHTML = String(week_traffic).replace(/(.)(?=(\d{3})+$)/g, '$1,');
  document.getElementById("avg_week_traffic").innerHTML = String(avg_week_traffic).replace(/(.)(?=(\d{3})+$)/g, '$1,') + " daily flights";
  document.getElementById("dif_week_prev_week").innerHTML = String(dif_week_prev_week) + '%';
  document.getElementById("dif_week_2019").innerHTML = String(dif_week_2019) + '%';
  document.getElementById("dif_week_prev_year").innerHTML = String(dif_week_prev_year) + '%';
  
  document.getElementById("y2d_traffic").innerHTML = String(y2d_traffic).replace(/(.)(?=(\d{3})+$)/g, '$1,');
  document.getElementById("avg_y2d_traffic").innerHTML = String(avg_y2d_traffic).replace(/(.)(?=(\d{3})+$)/g, '$1,') + " daily flights";
  document.getElementById("dif_y2d_2019").innerHTML = String(dif_y2d_2019) + '%';
  document.getElementById("dif_y2d_prev_year").innerHTML = String(dif_y2d_prev_year) + '%';

  }
  
</script>


```{r, echo=FALSE, message=FALSE, warning=FALSE}

nw_all_data <- read_sheet("https://docs.google.com/spreadsheets/d/14Lt3C3PCxxYCLXXYxRrFXd1s0tVZ3EQ8q9YfuYvx3Ww/edit?usp=sharing", sheet = "NW_Flights_All")


nw_all_data <- nw_all_data %>%
  # mutate_at(vars(FLIGHT_DATE), ~ as.Date(.x)) %>%
           mutate_at(vars(AVG_ROLLING_WEEK,AVG_ROLLING_WEEK_PREV_YEAR, AVG_ROLLING_WEEK_2020, AVG_ROLLING_WEEK_2019),round,0) 

# build plot
nw_all_data %>%
  plot_ly(width = 550, height = 370,
    x      = ~FLIGHT_DATE,
    y      = ~AVG_ROLLING_WEEK,
    name = '2022',
    type = 'scatter',
    mode = 'lines',
    line = list(color = '#003f5c', width = 3)
    ) %>%
  add_trace(
    y = ~AVG_ROLLING_WEEK_PREV_YEAR, 
    name = '2021',
    mode = 'lines',
    line = list(color = '#58508d', width = 1) 
    ) %>%
  add_trace(
    y = ~AVG_ROLLING_WEEK_2020, 
    name = '2020',
    mode = 'lines',
    line = list(color = '#bc5090', width = 1) 
    ) %>%
  add_trace(
    y = ~AVG_ROLLING_WEEK_2019, 
    name = '2019',
    mode = 'lines',
    line = list(color = '#ffa600', width = 2) 
    )%>%
  layout(
    autosize = F,
    title = list(text = '<b>Evolution of daily flights in the EUROCONTROL area</b>',
                 font = list(family = "Roboto", size = 16)),
    annotations = list(text = '(7-day moving average)', x = 0.5, y = 1.12,
                       showarrow = F, xref='paper', yref='paper',
                       xanchor='center', yanchor='auto',
                       font = list(family = 'Roboto', size = 14)),
    xaxis = list(showgrid = T, zeroline = T, showline = T, title = '',
                 tick0 = "2022-01-31", dtick = "M1", tickformat="%d %b") ,
    yaxis = list(showgrid = T, zeroline = T, showline = T, title = '',
                 hoverformat = ".0f", range = c(0, 41000), dtick = 10000,
                 constraintoward = 'bottom'),
    legend = list(orientation = 'h', xanchor = "center", x = 0.5, y=-0.2),
    hovermode = "x unified",
    margin = list(t=60)
    )%>%
    config(displaylogo = FALSE, modeBarButtons = list(list("toImage"))
           )
```
</div>
</div></div>
<br>

<!-- Stakeholders section -->
<div class="container" style="margin: 0px; padding: 0px">
#### Local/Stakeholder View
This section shows the top 40 aircraft operators, airports and country pairs in the Network in terms of flights (note that you can scroll down to see the full list). It also shows a list of all ANSPs ranked by total flights (including overflights). All rankings are presented for the latest available day, rolling week and year to date with a comparison with relevant previous periods.
(download files to be discussed)
<br>

<div class="flex-container" style=" margin-bottom: 5px;">
<div class="Mycolumn" >
<div class="col-12" style=" margin: 0px; " >
<h5>Aircraft Operators</h5>
<div class="resp-container">
<iframe class="resp-iframe" src="https://datastudio.google.com/embed/reporting/7cb12f8c-175b-4a00-8e41-4995bc78464c/page/p_01akhlhpsc" frameborder="0" style="border:0" allowfullscreen></iframe>
</div>
</div>
<br>
Aircraft operators in the table above are grouped as per levels 1 and 0 as defined in the <a href="https://www.eurocontrol.int/directory/airline-groups-lookup" target="_blank" >EUROCONTROL airline groups lookup</a>.
</div>
<div class="Mycolumn2">
<div class="col-12" style=" margin: 0px; " >
<h5>Airports</h5>
<div class="resp-container">
<iframe class="resp-iframe" src="https://datastudio.google.com/embed/reporting/71726332-25f1-412b-a902-50182fbc00c7/page/obGnC" frameborder="0" style="border:0" allowfullscreen></iframe>
</div></div>
</div>
</div>

<div class="flex-container" style=" margin-bottom: 5px;">
<div class="Mycolumn" >
<div class="col-12" style=" margin: 0px; " >
<h5>ANSPs</h5>
<div class="resp-container">
(TB added)</div>
</div>
</div>
<div class="Mycolumn2">
<div class="col-12" style=" margin: 0px; " >
<h5>Country Pairs</h5>
<div class="resp-container">
<iframe class="resp-iframe" src="https://datastudio.google.com/embed/reporting/74850727-d77f-4508-b312-114d40a3a7e6/page/tAPnC" frameborder="0" style="border:0" allowfullscreen></iframe>
</div>
</div>
</div></div>
</div>


