---
title: IATA delay codes
slug: iata-delay-codes
bibliography: [../../static/bib/aviation.bib]
link-citations: true
biblio-style: apalike
---
```{r echo = FALSE, message = FALSE, warning = FALSE}
# run setup script
source(here::here("_common.R"))
```


The following IATA delay codes (from "AHM 780 Aircraft Movement Message"
[@wiki:iatadelaycodes; @iata:delaycodes]
specification") standardise the reporting of commercial flight departure delays
by airlines.

```{r}
library(readr)
library(gt)
library(here)
library(dplyr)
library(glue)

dc <- read_csv("https://www.eurocontrol.int/performance/data/download/other/iata_delay_codes.csv")

dc %>% 
  mutate(ii = if_else(!is.na(cd), glue("{id} ({cd})"), glue("{id} (  )"))) %>% 
  group_by(group) %>% 
  gt(rowname_col = "ii") %>% 
   tab_header(
    title = "IATA delay codes"
  ) %>% 
  fmt_missing(
    columns = 2,
    missing_text = "..."
  ) %>% 
   cols_width(
    ii ~ px(80),
    id ~ px(30),
    cd ~ px(40)
  ) %>% 
    cols_label(
    id = "ID",
    cd = "Code",
    description = "Description"
  ) %>% 
    tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_row_groups()
  )
```


## See Also

* [IATA Delay Code 89](/definition/iata-code-89/)
* IATA Delay codes table ([PDF](/library/iata-delay-codes.pdf))
* Appendix of [All-causes delay and cancellations to Air Transport Europe Quaterly Reports](https://www.eurocontrol.int/library/search?keywords=All-causes+delay+and+cancellations+to+Air+Transport+Europe&sort_by=search_api_relevance)

## References

